---
title: "check.norm"
author: "Darren Thornbrugh"
date: "August 31, 2016"
output: html_document
---

## Article title: A practitioner’s guide for exploring water quality patterns using Principal Components Analysis and Procrustes
## Journal name: Environmental Monitoring and Assessment
## Author names: C. J. Sergeant, E. N. Starkey, K. K. Bartz, M. H. Wilson, F. J. Mueter
## Corresponding author: C. J. Sergeant, National Park Service, christopher_sergeant@nps.gov

```{r}

## Efficiently checking for normality in a range of parameter values using 'check.norm' function
## Written by Franz Mueter/Chris Sergeant
## Updated 18 September 2015

###################################
## Two easy ways to read-in data ##
###################################

# 1) choose your file from a Windows Explorer dialogue box that pops up after running code
# Note that 'header=T' assumes you have headers for each data column

wq = read.csv(file.choose(), header=T)

# 2) read-in data that is actively copied in your excel spreadsheet (.xls or .csv files)
# Copy your data in the spreadsheet, and while the marquee is still moving, run the code below
# Note that 'header=T' assumes you have headers for each data column

# Here is a good link discussing how to use this approach in more detail:
# http://www.dummies.com/how-to/content/how-to-use-the-clipboard-to-copy-and-paste-data-in.html

wq = read.table(file = "clipboard", sep = "\t", header=T)

# Quickly review the content and dimensions of imported data

head(wq); dim(wq)

```
## Setup function


```{r, echo=FALSE}
###########################################################
## Checking for normality in a range of parameter values ##
###########################################################

check.norm <- function(x, ...)
{

# Examine distribution of data and check against normal distribution
# x    is a vector of data. Additional graphics parameters can be supplied
#
# The function creates:
# (1) a histogram with an empirical density estimate, 
# (2) a boxplot
# (3) a normal q-q plot, and 
# (4) a plot of the empirical cumulative density function 
#     with the corresponding normal cdf. 
# In addition, the function returns the results from the 
# Shapiro-Wilks test of normality (p < 0.05 interpreted as non-normal data)
#
# Written by Franz Mueter. Last modified September 6, 2010
#

  par(mfrow=c(2,2))
  if(sum(is.na(x)) > 0)
  warning("NA's were removed before plotting")
  x <- x[!is.na(x)]
  hist(x, main = "Histogram and non-\nparametric density estimate", prob = T)
  iqd <- summary(x)[5] - summary(x)[2]
  lines(density(x, width = 2 * iqd))
  boxplot(x, main = "Boxplot", ...)
  qqnorm(x)
  qqline(x)
  plot.ecdf(x, main="Empirical and normal cdf")
  LIM <- par("usr")
  y <- seq(LIM[1],LIM[2],length=100)
  lines(y, pnorm(y, mean(x), sqrt(var(x))))
  shapiro.test(x)
}
```


```{r}
# Check normality of parameter from dataset 'wq'
# Where 'ENTER PARAMETER HERE' is written below, replace with parameter's column name in your data table (e.g. DO)

##check.norm(wq$ENTER PARAMETER HERE)
source("CheckNorm.Rmd")

check.norm(wq$SC)
check.norm(wq$DO)
check.norm(wq$pH)
check.norm(wq$Temp)

```

# For users that are familiar with working directories in R, you may want to consider creating a separate .R file
# composed of only the code for the check.norm function. Then, if that file is in your working directory and named
# 'CheckNorm.R', it can be called at any time using: source("CheckNorm.R")

